Automating Verification of MPC Security: $\metaprot$ to $\minifed$

### 2023-07-11 popl24/automation.tex
checkpoint

Since $|\runs(\prog)|$ grows exponentially in the size of
$\flips(\prog)$, scalability to larger programs is a concern.  However
computation of $\runs$ is amenable to optimizations, especially
parallelization- rather than computing the truth table
``horizontally'', it can be computed vertically. We can rewrite
$\minifed$ programs $\prog$ to stratified Datalog programs, where
$\runs(\prog)$ is equivalent to set of least Herbrand models obtained
by representing different random tapes and input secrets as different
fact bases. Details of this rewrite are provided in Appendix
\ref{section-bruteforce}. The model for each fact base can be computed
independently and in parallel, with the potential to apply other
matrix optimizations developed in recent work on HPC for Datalog
\cite{sakama2017linear,aspis2018linear,nguyen2022enhancing,nguyen2021efficient}.

>>>>

Since $|\runs(\prog)|$ grows exponentially in the size of
$\flips(\prog)$, scalability to larger programs is a concern.  However
computation of $\runs$ is amenable to optimizations, especially
parallelization- rather than computing the truth table
``horizontally'', it can be computed ``vertically''. We can rewrite
$\minifed$ programs $\prog$ to stratified Datalog programs, where
$\runs(\prog)$ is equivalent to set of least Herbrand models obtained
by representing different random tapes and input secrets as different
fact bases. Details of this rewrite are provided in Appendix
\ref{section-bruteforce}. The model for each fact base can be computed
independently and in parallel, with the potential to apply other
matrix optimizations developed in recent work on HPC for Datalog
\cite{sakama2017linear,aspis2018linear,nguyen2022enhancing,nguyen2021efficient}.

### 2023-07-11 popl24/composition.tex
checkpoint

\begin{proof}
  Given well-typedness of $\prog_1;E[\mathit{gate}(\ttt{g},v_1,v_2)]$, we have for some $\ttt{"g1"}$ and $\ttt{"g2"}$:
  $$
  \begin{array}{lcl}
   v_1 &=& \ttt{\{c1 = v[1,"g1out"]; c2 = v[2,"g1out"]\}}\\
   v_2 &=& \ttt{\{c1 = v[1,"g2out"]; c2 = v[2,"g2out"]\}}\\
    v &=& \ttt{\{c1 = v[1,"gout"]; c2 = v[2,"gout"]\}}
  \end{array}
  $$
  Let $\prog$ be as defined in Definition \ref{definition-gmwgate-certification}.
  We observe:
  $$
  {\footnotesize
    \begin{array}{c}
      \prog_2 = \\
      \prog[\ttt{v[1,"g1out"]}/\ttt{flip[1,"s11"]}][\ttt{v[2,"g1out"]}/\ttt{flip[2,"s12"]}][\ttt{v[1,"g2out"]}/\ttt{flip[1,"s21"]}][\ttt{v[2,"g2out"]}/\ttt{flip[2,"s22"]}]
    \end{array}
  }
  $$
  Now, the assumption of well-typedness also assures that $\ttt{"g"}$
  has not been previously used as a gate identifier in $\prog_1$,
  since otherwise the views $\ttt{v[1,"gout"]}$ and $\ttt{v[2,"gout"]}$
  would have been previously defined. This ensures $\vars(\prog_1,x)
  \cap \vars(\prog) = \varnothing$ for:
  $$x \in \{ \ttt{v[1,"g1out"]}, \ttt{v[2,"g1out"]}, \ttt{v[1,"g2out"]}, \ttt{v[2,"g2out"]}\}$$
  given condition (iii) of Definition \ref{definition-gmwgate-certification}.
  Thus by Lemma \ref{lemma-substitution} we have:
  $$\sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[1,"g1out"],v[1,"g2out"],v[2,"g1out"],v[2,"g2out"]\}}}{\ttt{\{v[1,"gout"]\}}}$$
  and
  $$\sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[1,"g1out"],v[1,"g2out"],v[2,"g1out"],v[2,"g2out"]\}}}{\ttt{\{v[2,"gout"]\}}}$$
  Now, by Lemmas \ref{lemma-noninterference} and \ref{lemma-separation} and condition (iii) of Definition
  \ref{definition-gmwgate-certification} we then have:
  $$\sep{\progtt(\prog_1;\prog_2)}{(S_{\{1\}} \cup V_{\{2\}})}{\ttt{\{v[2,"gout"]\}}} \quad \text{and} \quad
  \sep{\progtt(\prog_1;\prog_2)}{(S_{\{2\}} \cup V_{\{1\}})}{\ttt{\{v[1,"gout"]\}}}$$
  so by precondition assumptions and Lemma \ref{lemma-separation} we have:
  $$\sep{\progtt(\prog_1;\prog_2)}{S_{\{1\}}}{V'_{\{2\}}} \quad \text{and} \quad
  \sep{\progtt(\prog_1;\prog_2)}{S_{\{2\}}}{V'_{\{1\}}}$$
\end{proof}

Here we show that encoding also preserves the desired invariant.
The result here is more straightforward than it is for gates, because
the variables used during encoding are guaranteed to be distinct from
the rest of the program.
\begin{restatable}[GMW Encode Invariant]{lemma}{gmwencode}
  \label{lemma-gmw-encode}
  Given well-typed $\prog_1;E[\mathit{encode}\ttt{(s1,s2)}]$ for certified $\mathit{encode}$, let:
  $$
  \config{\prog_1}{E[\mathit{encode}(\ttt{s1},\ttt{s2})]} \redxs
  \config{\prog_1;\prog_2}{E[v]}
  $$
  If we have the following preconditions, where $\iov(\prog_1) = S' \cup V$:
  \begin{enumerate}[\hspace{5mm}i.]
  \item $\sep{\progd(\prog_1)}{S_{\{2\}}}{V_{\{1\}}}$
  \item $\sep{\progd(\prog_1)}{S_{\{1\}}}{V_{\{2\}}}$
  \end{enumerate}
  then we have as a postconditions, where $\iov(\prog_1;\prog_2) = S' \cup V'$: 
  \begin{enumerate}[\hspace{5mm}i.]
  \item $\sep{\progd(\prog_1;\prog_2)}{S'_{\{2\}}}{V'_{\{1\}}}$
  \item $\sep{\progd(\prog_1;\prog_2)}{S'_{\{1\}}}{V'_{\{2\}}}$
  \end{enumerate}
\end{restatable}

>>>>

\begin{proof}
  Given well-typedness of $\prog_1;E[\mathit{gate}(\ttt{"g"},v_1,v_2)]$, we have for some $\ttt{"g1"}$ and $\ttt{"g2"}$:
  $$
  \begin{array}{lcl}
   v_1 &=& \ttt{\{c1 = v[1,"g1out"]; c2 = v[2,"g1out"]\}}\\
   v_2 &=& \ttt{\{c1 = v[1,"g2out"]; c2 = v[2,"g2out"]\}}\\
    v &=& \ttt{\{c1 = v[1,"gout"]; c2 = v[2,"gout"]\}}
  \end{array}
  $$
  Let $\prog$ be as defined in Definition \ref{definition-gmwgate-certification}.
  We observe:
  $$
    \begin{array}{lcl}
      \prog_2 &=&
           \prog[\ttt{v[1,"g1out"]}/\ttt{flip[1,"s11"]}][\ttt{v[2,"g1out"]}/\ttt{flip[2,"s12"]}]\\
      & & \phantom{\prog}[\ttt{v[1,"g2out"]}/\ttt{flip[1,"s21"]}][\ttt{v[2,"g2out"]}/\ttt{flip[2,"s22"]}]
    \end{array}
  $$
  Now, the assumption of well-typedness also assures that $\ttt{"g"}$
  has not been previously used as a gate identifier in $\prog_1$,
  since otherwise the views $\ttt{v[1,"gout"]}$ and $\ttt{v[2,"gout"]}$
  would have been previously defined. This ensures $\vars(\prog_1,x)
  \cap \vars(\prog) = \varnothing$ for:
  $$x \in \{ \ttt{v[1,"g1out"]}, \ttt{v[2,"g1out"]}, \ttt{v[1,"g2out"]}, \ttt{v[2,"g2out"]}\}$$
  given condition (iii) of Definition \ref{definition-gmwgate-certification}.
  Thus by Lemma \ref{lemma-substitution} we have:
  $$\sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[1,"g1out"],v[1,"g2out"],v[2,"g1out"],v[2,"g2out"]\}}}{\ttt{\{v[1,"gout"]\}}}$$
  and
  $$\sep{\progtt(\prog_1;\prog_2)}{\ttt{\{v[1,"g1out"],v[1,"g2out"],v[2,"g1out"],v[2,"g2out"]\}}}{\ttt{\{v[2,"gout"]\}}}$$
  Now, by Lemmas \ref{lemma-noninterference} and \ref{lemma-separation} and condition (iii) of Definition
  \ref{definition-gmwgate-certification} we then have:
  $$\sep{\progtt(\prog_1;\prog_2)}{(S_{\{1\}} \cup V_{\{2\}})}{\ttt{\{v[2,"gout"]\}}} \quad \text{and} \quad
  \sep{\progtt(\prog_1;\prog_2)}{(S_{\{2\}} \cup V_{\{1\}})}{\ttt{\{v[1,"gout"]\}}}$$
  so by precondition assumptions and Lemma \ref{lemma-separation} we have:
  $$\sep{\progtt(\prog_1;\prog_2)}{S_{\{1\}}}{V'_{\{2\}}} \quad \text{and} \quad
  \sep{\progtt(\prog_1;\prog_2)}{S_{\{2\}}}{V'_{\{1\}}}$$
\end{proof}

Here we show that encoding also preserves the desired invariant.
The result here is more straightforward than it is for gates, because
the variables used during encoding are guaranteed to be distinct from
the rest of the program.
\begin{restatable}[GMW Encode Invariant]{lemma}{gmwencode}
  \label{lemma-gmw-encode}
  Given well-typed $\prog_1;E[\mathit{encode}\ttt{("s1","s2")}]$ for certified $\mathit{encode}$, let:
  $$
  \config{\prog_1}{E[\mathit{encode}(\ttt{"s1"},\ttt{"s2"})]} \redxs
  \config{\prog_1;\prog_2}{E[v]}
  $$
  If we have the following preconditions, where $\iov(\prog_1) = S' \cup V$:
  \begin{enumerate}[\hspace{5mm}i.]
  \item $\sep{\progd(\prog_1)}{S_{\{2\}}}{V_{\{1\}}}$
  \item $\sep{\progd(\prog_1)}{S_{\{1\}}}{V_{\{2\}}}$
  \end{enumerate}
  then we have as a postconditions, where $\iov(\prog_1;\prog_2) = S' \cup V'$: 
  \begin{enumerate}[\hspace{5mm}i.]
  \item $\sep{\progd(\prog_1;\prog_2)}{S'_{\{2\}}}{V'_{\{1\}}}$
  \item $\sep{\progd(\prog_1;\prog_2)}{S'_{\{1\}}}{V'_{\{2\}}}$
  \end{enumerate}
\end{restatable}
